---
title: "RNA-seq_DE-analysis"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## GitHub Documents

This is an R Markdown format used for publishing markdown documents to GitHub. When you click the **Knit** button all R code chunks are run and a markdown file (.md) suitable for publishing to GitHub is generated.
## Library the packages
```{r echo=FALSE}
library("openxlsx")
library("limma")
library("DESeq2")
library("edgeR")
library("ggplot2")
library("Cairo")
library("dplyr")
library(VennDiagram)
```


## Data Input: RNA-seq raw datas

```{r}
rna_seq_data0 <- read.xlsx(file.choose())
```
## Data Input: GeneName
```{r}
Genename <- read.xlsx(file.choose())
```


## Data handling
```{r echo=FALSE}
rna_seq_data <- merge(rna_seq_data0,Genename,by.x="gene_id",by.y="ENTREZ_GENE_ID",all.x = TRUE)
sum(is.na(rna_seq_data$symbol))
colnamber<-ncol(rna_seq_data)-1
colnamber
countdata<- as.matrix(floor(rna_seq_data[1:nrow(rna_seq_data),4:colnamber]))
countdata <- subset(countdata,select=c(1,2,5,6))
condition <- as.factor(c(rep("sh1",2),rep("Ctrl",2)))
col_data <-data.frame(row.names = colnames(countdata),
                      group_list=condition)
```

## DESeq2
```{r}
dds <- DESeqDataSetFromMatrix(countData = countdata,
                              colData = col_data,
                              design = ~group_list)

rownames(dds)<-rna_seq_data$gene_id
dds2 <- DESeq(dds)
rld <- rlogTransformation(dds2)
exprSet = assay(rld)

boxplot(countdata)
boxplot(exprSet)
```

## DESeq2 pvalue
```{r}
resultsNames(dds2)
res <- results(dds2)
res_ordered <- as.data.frame(res[order(res$padj),])
head(res$padj)
```

## DESeq2 log2Foldchange
```{r}
head(res$log2FoldChange)
```

## volcanoplot data handling
```{r}
res_ordered$threshold <- as.factor(ifelse(res_ordered$pvalue<0.05&abs(res_ordered$log2FoldChange)>=1,
                                            ifelse(res_ordered$log2FoldChange>1,"Up","Down"),"Not"))

downname <- head(subset(res_ordered,threshold=="Down"),13)  #subset the important gene

name <- subset(rna_seq_data,gene_id==rownames(downname))

sum(is.na(rna_seq_data$symbol))
upname <- head(subset(res_ordered,threshold=="Up"),13)  #subset the important gene

dename <- subset(res_ordered,pvalue<0.05&threshold!="Not")
```

## volcano plot
```{r}
ggplot(res_ordered,
       aes(x=res_ordered$log2FoldChange,
           y=-log10(res_ordered$pvalue),
           colour = threshold))+
  scale_color_manual(values=c("skyblue","grey","red"))+
  geom_point(size = 2.5)+
  xlim(c(-4,4))+
  ylab("-log10 p-value")+
  xlab("log2 fold change")+
  labs(title="Volcano of CBSsh")+
  geom_vline(xintercept = c(-1,1),lty=4,col="black",lwd=0.6)+
  geom_hline(yintercept = -log10(0.05),lty=4,col="black",lwd=0.6)+
  theme(legend.position = "right",
        panel.grid = element_blank(),
        legend.title = element_blank(),
        legend.text = element_text(face = "bold",color = "black",size = 12),
        plot.title = element_text(hjust = 0.5))+
  annotate(geom ="text",
           x=downname$log2FoldChange,
           y=-log10(downname$pvalue),
           label=rownames(downname),
           size = 2.3)+
  annotate(geom ="text",
           x=upname$log2FoldChange,
           y=-log10(upname$pvalue),
           label=rownames(upname),
           size = 2.3)

```

## Data Input 
```{r}
venn1 <- read.xlsx(file.choose())
venn2 <- read.xlsx(file.choose())
venn3 <- read.xlsx(file.choose())
```

## venndiagram
```{r}
venn.diagram(list(list1 =venn1$gene_id, list2=venn2$geneid),
             filename = "1_1.tif",
             col = "transparent",
             fill = c("cornflowerblue","darkorchid1"),
             label.col = "black",
             cat.col = c("blue","red"),
             cat.dist = c(0.03,0.03),
             cat.pos = c(-8,8),
             cat.cex = 1.5,
             cex = 2,
             main = "venn plot of DE genes",
             main.cex = 2)
```

## GO-analysis
```{r}
library("DOSE")
library("clusterProfiler")
ego3 <- enrichGO(gene = c$GENE.ID,
                 OrgDb = "org.Mm.eg.db",
                 ont = "CC",
                 pvalueCutoff = 0.05,
                 readable = TRUE)
summary(as.data.frame(ego))
```

## KEGG-analysis
```{r}
ekk <- enrichKEGG(gene =c$GENE.ID,
                  organism = "mmu",
                  pvalueCutoff = 0.05,
                  pAdjustMethod = "BH",
                  use_internal_data = TRUE)
```

##Pathview
```{r}
require("pathview")
pathview(gene.data = c[,2],
         gene.idtype = "entrez",
         pathway.id = ekk,
         species = "mmu",
         # limit = list(gene = max(log2(LH_data$`Act.HSC.vs.qHSC.(Fold)`)),cpd = 1),
         bins = list(gene = 10,cpd=10),
         out.suffix = "LH_pathview_1",
         kegg.native = TRUE)
```






